### 2022-02-23 manuscript/figure_chunks.Rmd

```{r fig-consort, fig.width = 7.086614166, fig.height = 5.905511805, fig.cap = 'Analysis inclusion scheme.'}

paper_figures$consort$plot

```

__Figure \@ref(fig:fig-consort). Analysis inclusion scheme.__

```{r fig-cyto-screen, fig.width = 7.086614166, fig.height = 4.724409444, fig.cap = 'Screening of the flow cytometry features for the differences between healthy controls, moderate and severe COVID-19.'}

paper_figures$cyto_screen$plot

```

__Figure \@ref(fig:fig-cyto-screen). Screening of the flow cytometry features for the differences between healthy controls, moderate and severe COVID-19.__

```{r fig-subset-distribution, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Monocyte subset distribution in healthy controls, moderate and severe COVID-19.'}

paper_figures$subset_distribution$plot

```

__Figure \@ref(fig:fig-subset-distribution). Monocyte subset distribution in healthy controls, moderate and severe COVID-19.__

```{r fig-clustering, fig.width = 7.086614166, fig.height = 5.905511805, fig.cap = 'Unsupervised clustering of healthy controls and COVID-19 individuals by myeloid leukocyte cytometry features.'}

paper_figures$clustering$plot

```

__Figure \@ref(fig:fig-clustering). Unsupervised clustering of healthy controls and COVID-19 individuals by myeloid leukocyte cytometry features.__

```{r fig-clust-inflammation, fig.width = 7.086614166, fig.height = 8.661417314, fig.cap = 'Differences in systemic inflammation status between the clusters of study participants.'}

paper_figures$clust_inflammation$plot

```

__Figure \@ref(fig:fig-clust-inflammation). Differences in systemic inflammation status between the clusters of study participants.__